A partisan affair: Mapping patronage in municipal bureaucracies of Brazil

### 2021-03-10 paper/draft.tex
add image
add section on why and who benefits from patronage

\begin{figure}[H]
    \centering
    \includegraphics[width = 10cm, height = 10cm]{figures/partisanship/plot_partisan_by_year.pdf}
    \caption{\textbf{Number of active party members by year}. There has been a rapid growth in the number of party members in the period between 1997 and 2020.}
    \label{fig:map_pooled}
\end{figure}

@@

\subsection{Who benefits from patronage?}

Who should politicians buy off: rich or poor voters? There are competing views on who benefits from patronage. On the one hand, from a politician's perspective it would be cheaper to buy voters who are poor, insofar as the return on investment would be greater for poorer voters, who need and value more monetary incentives \citet{stokes2013brokers}. Other scholars argue that in order to maximize electoral returns, politicians target swing voters, those who are indifferent to either party and therefore, again, require the least resources to buy off \citet{dixit1996determinants}.

On the other, there are good reasons to expect politicians to target rich voters as well. There is a well-established literature on money and politics,

>>>>

\begin{figure}[H]
    \centering
    \includegraphics[width = 10cm, height = 8cm]{figures/partisanship/plot_partisan_by_year.pdf}
    \caption{\textbf{Number of active party members by year}. There has been a rapid growth in the number of party members in the period between 1997 and 2020.}
    \label{fig:map_pooled}
\end{figure}

@@

\subsection{Why engage in patronage?}

Why do politicians engage in patronage? And who should politicians buy off? There are competing views on who benefits from patronage, and what is motivating it in the first place. On the one hand, from a politician's perspective it can be cheaper to buy voters who are poor, insofar as these voters need and value more monetary incentives \citet{stokes2013brokers}. Other scholars argue that in order to maximize electoral returns, politicians target swing voters, those who are indifferent to either party and therefore, again, require the least resources to buy off \citet{dixit1996determinants}.

The socioeconomic status of voters is not, however, the only consideration that politicians take into place. Patronage can be used by politicians to reward political loyalists for resources they have contributed for an electoral campaign, a practice generally denoted as \emph{prebendism}\citep{van2007meet}. In Brazil, studies have found that campaign contributors are more likely to be employed in the local bureaucracy, in the magnitude of a 30 percent increase in the probability of being hired \citep{colonnelli2018patronage}. Additionally, there is strong evidence that employment in the bureaucracy is used as a spoil to reward party loyalists, again with evidence from Brazilian local bureaucracies \citep{brollo2017victor}.

Why should patronage occur post-election? There is an important distinction to be made between different forms of clientelism such as disbursement of campaign materials \citep{stokes2005perverse,nichter2008vote} and another form of clientelism which can only occur after a politician takes office. In particular, a politician cannot offer job to a voter prior to holding office. Moreover, public sector jobs are coarse goods in the sense that they cannot be infinitely divided among voters. There are tight budget and administrative constraints that distinguishes a public sector jobs from other forms of resources available to politicians.

If public sector jobs are scarce, and there are only so many spots that can be filled post-election, a different form of consideration takes place from the standpoint of the politician. Differently from an electoral race in which each voter theoretically contributes the same exact amount to the race, a single vote, constituencies can vary widely with respect to the resources that they bring in to the table. A well established literature on the influence of money on politics has documented how campaign contributions play a crucial role in electoral outcomes in both the developed and developing world \citep{claessens2008political}. What is overseen is the fact that patronage can lock-in wealthy patrons into the bureaucracy and incentivize greater campaign contributions in the future, effectively securing important revenue streams for both the politician and her party \citep{robinson2013political}.

This is in fact what the data shows.
